\bfseries A simple momentum-transport model for sailing

### 2021-07-11 paper/main.tex
numbers

% to-do items:
% ------------
% - Put (m/s) onto hourglass figure captions.
% - Make figure comparing different boats.
% - Change citations to Tao to make it clear that the model is wrong!!
% - Make sure I cite {symmetry} appropriately in various places. Probably also point out that it is a fun longer read.
% - Make sure we aren't claiming that the model is NOVEL. It is just SIMPLE.
% - Reconsider the post-equation commas? (HWR SAY).
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?
% - Do we still mention "tensors" anywhere? make sure the abstract is consistent with the body.

@@

We could define \emph{GoodSailing(tm)} to be setting the sail orientation optimally, given the orientation of the boat.
This might look something like the following:
\begin{align}\label{eq:good}
    \uvec_{\perp\sail}^\good &\leftarrow \argmax_{\uvec_{\perp\sail}} \left[\uvec_{\parallel\keel}^\top\,(\vboat-\vwater)\right] ~,
\end{align}
where $\uvec_{\perp\sail}^\good$ is the best orientation of the (planar) sail conditioned on the orientation of the keel,
$\uvec_{\parallel\keel}$ is the unit vector pointing along the keel (parallel to the keel, in the forwards direction),
the inner product gives the projection of the boat--water velocity vector onto the boat heading direction,
and (for optimization purposes) the velocity difference $\vboat-\vwater$ is implicitly a function of the sail and keel orientations, according to the model of \secref{sec:steady}.
This optimization is one-dimensional, but it is a global optimization, so it requires some attention; details of our implmentation of the optimization are given in \secref{sec:implementation}.
Some examples of GoodSailing(tm) settings of sail and keel are shown in \figref{fig:good}, along with the resulting boat velocity vectors.
A summary of all possible boat--water velocity vectors possible, for GoodSailing(tm) at many different keel orientations, is shown in \figref{fig:hourgood}.
This latter figure is the HOGG WHAT DIAGRAM that is standard in all books on the subject of sailing (HOGG GIVE REFERENCES).
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{good.pdf}
  \caption{Visualizations of GoodSailing(tm) for different settings of the orientation of the keel.
  Conventions are the same as in \figref{fig:steady}.
  GoodSailing(tm) is unsuccessful in the last two cases, where the pointing is too close to directly upwind.
  In these cases the the maximum of the GoodSailing(tm) objective \eqref{eq:good} is negative.\label{fig:good}}
  \figurerule
\end{figure}
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-good.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for GoodSailing(tm) for a set of randomly oriented keels.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:good}.
  Compare this to \figref{fig:hourglass}.
  As is visible in \figref{fig:good}, when the keel orientation is too close to directly upwind, the boat goes downwind instead; there are no \emph{directly} upwind options. This figure is directly comparable to [HOGG WHAT FIGURES IN THE BOOKS ABOUT SAILING?].\label{fig:hourgood}}
  \figurerule
\end{figure}

@@

We show examples of BetterSailing(tm) in \figref{fig:better}, where we show the direction to the destination and the settings of the keel and sail.
In these examples, we have set the velocity of the destination equal to the velocity of the water, but this will not be the case in general.
In \figref{fig:hourbetter} we show the velocities obtained in BetterSailing(tm), which are a subset of the velocities obtained in GoodSailing(tm).
In particular, the velocities that are not optimal for some particular destination are not used.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{better.pdf}
  \caption{Visualizations of BetterSailing(tm) for different directions to a distant destination.
  Conventions are the same as in \figref{fig:steady} and \figref{fig:good}, except that now the destination direction is indicated with a dashed path line.
  These plots assume that the destination is stationary with respect to the water ($\vdest-\vwater=\vec{0}$).
  BetterSailing(tm) often involves sailing at a substantial angle to the direct line towards the destination.
  In these cases, the idea is that the best route to the destination involves frequent tacking back and forth across the path to the destination.\label{fig:better}}
  \figurerule
\end{figure}
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-better.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for BetterSailing(tm) for a set of randomly oriented destination vectors.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:better}.
  Compare this to \figref{fig:hourglass} and also to \figref{fig:hourgood}.\label{fig:hourbetter}}
  \figurerule
\end{figure}

@@

In all of the experiments and figures given above, we made use of a model boat with the following effective areas:
\begin{align}
  A_\sail &= XXX\,\mm \\
  A_\keel &= YYY\,\mm \\
  A_\above &= ZZZ\,\mm \\
  A_\below &= QQQ\,\mm ~.
\end{align}
This boat is designed such that $\rho_\air\,A_\sail = \rho_\water\,A_\keel$, which seems sensible, but is perhaps not industry standard.
This boat also has very good drag ratios or lift ratios:
The ratio of effective area of sail to hull-above is 256, as is the ratio of effective area of keel to hull-below.
All these ratios were set not by analyzing any real boats, but by making the velocity curve shown in \figref{fig:hourgood} look reasonable, or accord with our intuitions and expectations.
HOGG: NEED A BETTER NAME THAN VELOCITY CURVE; WHAT IS THIS DIAGRAM CALLED?
Recall, for one, that our boat is a brutal approximation!
For two, these areas are \emph{effective areas} within this brutal approximation, not measured areas.
Probably the least realistic of these effective-area ratios is $A_\keel/A_\below$; probably real boats rarely have this keel lift ratio in the hundreds.
In this sense, maybe our model is more appropriate for an ice yacht than a sailboat?

>>>>

% to-do items:
% ------------
% - Make figure comparing different boats.
% - Change citations to Tao to make it clear that the model is wrong!!
% - Make sure I cite {symmetry} appropriately in various places. Probably also point out that it is a fun longer read.
% - Make sure we aren't claiming that the model is NOVEL. It is just SIMPLE.
% - Reconsider the post-equation commas? (HWR SAY).
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?
% - Do we still mention "tensors" anywhere? make sure the abstract is consistent with the body.
% - What claims in the literature exist about maximum boat speeds? Have I seen a 2x wind claim? We can beat that. Have I seen a "never faster downwind than the wind" or than 0.5 times the wind claim? We can beat that too.

@@

We could define \emph{GoodSailing(tm)} to be setting the sail orientation optimally, given the orientation of the boat.
This might look something like the following:
\begin{align}\label{eq:good}
    \uvec_{\perp\sail}^\good &\leftarrow \argmax_{\uvec_{\perp\sail}} \left[\uvec_{\parallel\keel}^\top\,(\vboat-\vwater)\right] ~,
\end{align}
where $\uvec_{\perp\sail}^\good$ is the best orientation of the (planar) sail conditioned on the orientation of the keel,
$\uvec_{\parallel\keel}$ is the unit vector pointing along the keel (parallel to the keel, in the forwards direction),
the inner product gives the projection of the boat--water velocity vector onto the boat heading direction,
and (for optimization purposes) the velocity difference $\vboat-\vwater$ is implicitly a function of the sail and keel orientations, according to the model of \secref{sec:steady}.
This optimization is one-dimensional, but it is a global optimization, so it requires some attention; details of our implmentation of the optimization are given in \secref{sec:implementation}.
Some examples of GoodSailing(tm) settings of sail and keel are shown in \figref{fig:good}, along with the resulting boat velocity vectors.
A summary of all possible boat--water velocity vectors possible, for GoodSailing(tm) at many different keel orientations, is shown in \figref{fig:hourgood}.
This latter figure is the HOGG WHAT DIAGRAM that is standard in all books on the subject of sailing (HOGG GIVE REFERENCES).
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{good.pdf}
  \caption{Visualizations of GoodSailing(tm) for different settings of the orientation of the keel.
  Conventions are the same as in \figref{fig:steady}.
  GoodSailing(tm) is unsuccessful in the last two cases, where the pointing is too close to directly upwind.
  In these cases the the maximum of the GoodSailing(tm) objective \eqref{eq:good} is negative.\label{fig:good}}
  \figurerule
\end{figure}
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-good.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for GoodSailing(tm) for a set keel orientations.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:good}.
  Compare this to \figref{fig:hourglass}.
  As is visible in \figref{fig:good}, when the keel orientation is too close to directly upwind, the boat goes downwind instead; there are no \emph{directly} upwind options. This figure is directly comparable to [HOGG WHAT FIGURES IN THE BOOKS ABOUT SAILING?].\label{fig:hourgood}}
  \figurerule
\end{figure}

@@

We show examples of BetterSailing(tm) in \figref{fig:better}, where we show the direction to the destination and the settings of the keel and sail.
In these examples, we have set the velocity of the destination equal to the velocity of the water, but this will not be the case in general.
In \figref{fig:hourbetter} we show the velocities obtained in BetterSailing(tm), which are a subset of the velocities obtained in GoodSailing(tm).
In particular, the velocities that are not optimal for some particular destination are not used.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{better.pdf}
  \caption{Visualizations of BetterSailing(tm) for different directions to a distant destination.
  Conventions are the same as in \figref{fig:steady} and \figref{fig:good}, except that now the destination direction is indicated with a dashed path line.
  These plots assume that the destination is stationary with respect to the water ($\vdest-\vwater=\vec{0}$).
  BetterSailing(tm) often involves sailing at a substantial angle to the direct line towards the destination.
  In these cases, the idea is that the best route to the destination involves frequent tacking back and forth across the path to the destination.\label{fig:better}}
  \figurerule
\end{figure}
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-better.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for BetterSailing(tm) for a set of destination vectors of different orientations.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:better}.
  Compare this to \figref{fig:hourglass} and also to \figref{fig:hourgood}.\label{fig:hourbetter}}
  \figurerule
\end{figure}

@@

In all of the experiments and figures given above, we made use of a model boat with the following effective areas:
\begin{align}
  A_\sail &= 1400\,\mm \\
  A_\keel &= 2\,\mm \\
  A_\above &= 5.6\,\mm \\
  A_\below &= 0.008\,\mm ~.
\end{align}
This boat is designed such that $\rho_\air\,A_\sail = \rho_\water\,A_\keel$, which seems sensible, but is perhaps not industry standard.
This boat also has very good drag ratios or lift ratios:
The ratio of effective area of sail to hull-above is 250, as is the ratio of effective area of keel to hull-below.
All these ratios were set not by analyzing any real boats, but by making the velocity curve shown in \figref{fig:hourgood} look reasonable, or accord with our intuitions and expectations.
HOGG: NEED A BETTER NAME THAN VELOCITY CURVE; WHAT IS THIS DIAGRAM CALLED?
Recall, for one, that our boat is a brutal approximation!
For two, these areas are \emph{effective areas} within this brutal approximation, not measured areas.
Probably the least realistic of these effective-area ratios is $A_\keel/A_\below$; probably real boats rarely have this keel lift ratio in the hundreds.
In this sense, maybe our model is more appropriate for an ice yacht than a sailboat?
